Evolutionary Algorithm for the Bonded Molecular Conformation Problem

### 2018-08-31 documentation/paper.tex
Added tables and figures

\begin{document}
\maketitle
\tableofcontents
\pagebreak

@@

Finally, $\alpha_i$ is locally optimised by applying BFGS with this single angle
as its only optimisation parameter, with the rest of the angles fixed. Energy
and gradients for use in the BFGS algorithm are computed using the entire
molecule, while only the subset is optimised, leaving the rest of the angles
unchanged. This optimisation of the single crossover point effectively locally
optimises the relative orientation of the two connected segments.
%TODO: Should I leave this paragraph in?
Crossover operators with multiple crossover-points were not found to be
beneficial during experimentation and were even detrimental in some cases. As
such, they have been excluded from this study. The focus is primarily on
evaluating the use of annealing as a mutation operator and therefore the
crossover operation has been kept simple.

@@

\subsubsection{Mutation Rate}

@@

@@

A random angle is selected and the current fitness and the current value of this
	angle is stored.The selected angle is set to a random value between $-\pi$
	and $\pi$ and the fitness is recomputed. If the new fitness is less than the
	previous fitness, revert the angle to its previous state. This process is
	then repeated until the maximum number of iterations is reached. This limit
	is important because after many generations it becomes increasingly
	difficult to find a random change in an angle that does not give lower
	energy, which would otherwise result in an infinite loop.

@@

\paragraph{Simulated Annealing Mutation}

@@

\subsection{Local Optimisation}

% The final stage of the algorithm, after a candidate configuration has been
% found by the GA and its immediate vicinity has been further explored through
% mutation, is to refine the configuration using gradient-based local search.
% This stage will use the gradient of the objective function in order to
% guarantee that the bottom of the current basin will be reached. The
% quasi-Newton method Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at
% this stage resulting in a fast descent towards the current local minimum at
% each stage.

% The coordinates of this locally optimised configuration will also be passed
% back to the GA. This is known as a `Lamarkian' process and has been found to
% be beneficial when used with this problem \cite{doi:10.1002/qua.560440214}.

@@

@@

\section{Discussion}

@@

>>>>

\begin{document}
\maketitle
% \tableofcontents
% \pagebreak

@@

Finally, $\alpha_i$ is locally optimised by applying BFGS with this single angle
as its only optimisation parameter, with the rest of the angles fixed. Energy
and gradients for use in the BFGS algorithm are computed using the entire
molecule, while only the subset is optimised, leaving the rest of the angles
unchanged. This optimisation of the single crossover point effectively locally
optimises the relative orientation of the two connected segments.

%TODO: Should I leave this paragraph in?
Crossover operators with multiple crossover-points were not found to be
beneficial during experimentation and were even detrimental in some cases. As
such, they have been excluded from this study. The focus is primarily on
evaluating the use of annealing as a mutation operator and therefore the
crossover operation has been kept simple.

@@

\subsubsection{Mutation Rate} \label{mutationrate}

@@

The following operators mutate a given configuration in both biased and
unbiased ways.

@@

A random angle is selected and both the current fitness of the molecule and the
current value of this angle is stored.The selected angle is set to a random
value between $-\pi$ and $\pi$ and the fitness is recomputed. If the new fitness
is worse than the previous fitness, the angle is reverted to its previous state.
This process is then repeated until either the maximum number of iterations is
reached, or an improved configuration is found. This limit is important because
after many generations it becomes increasingly difficult to find a random change
in an angle that does not give lower energy, which would otherwise result in an
infinite loop.

@@

This is an unbiased mutation operator which simply selects a random angle
$\alpha_i$ and sets it to a random new value between $-\pi$ and $\pi$. This
operator has no regard for whether the mutation improves or worsens state, and
primarily serves the purpose of reducing homogeneity and premature convergence
within the population.

\paragraph{Annealed Mutation}

@@

@@

The following GA parameters are used:
\begin{description}
	\item[Population Size] 12
	\item[Maximum Generations] 100
	\item[Mutation Rate] Deterministic and based on conditions as described in
	section~\ref{mutationrate}
\end{description}
The GA works by continually sorting a large pool (of size $n_{pop} \times 2$) of
configurations in order to rank them in order of fitness and subsequently select
states for crossover and mutation operators. The population is initialised with
random molecules and sorted and then the procedure outlined in REF are carried
out until either a known global optimum is found or the maximum number of
generations is reached.

%%%% GA Pseudocode here

The final stage of each GA iteration, after candidate configurations have been
found by the GA and its immediate vicinity has been further explored through
mutation, is to refine the configuration using gradient-based local search. This
stage uses the gradient of the objective function in order to guarantee that the
bottom of the current basin will be reached. The quasi-Newton method
Broyden–Fletcher–Goldfarb–Shanno (BFGS) will be applied at this stage resulting
in a fast descent towards the current local minimum at each stage. The
coordinates of this locally optimised configuration are also be passed back to
the GA so that improvements are retained in future iterations. This is known as
a `Lamarkian'process and has previously been found to be beneficial when used
with this problem \cite{doi:10.1002/qua.560440214}.

@@

\begin{figure}
	\includegraphics[width=\textwidth]{nvse.pdf}
	\caption{Lennard-Jones Potential as N size increases}
	\label{fig:nvse}
\end{figure}

\section{Conclusion}

@@

\begin{table}[]
	\label {tab:nosa}
	\centering
	\begin{tabular}{cccc|cccc}
	$N$  & Optimal & Found & Gen. & $N$  & Optimal & Found  & Gen. \\
	\hline
	3  & -3.0    & -3.0  & 0    & 30 & -77.3   & -77.2  & 89   \\
	4  & -5.1    & -5.1  & 1    & 31 & -79.5   & -79.6  & 70   \\
	5  & -7.2    & -7.2  & 1    & 32 & -82.9   & -82.9  & 88   \\
	6  & -9.3    & -9.4  & 52   & 33 & -86.1   & -86.2  & 87   \\
	7  & -12.5   & -12.5 & 67   & 34 & -88.4   & -88.5  & 99   \\
	8  & -14.7   & -14.7 & 75   & 35 & -91.7   & -90.7  & 99   \\
	9  & -16.9   & -16.9 & 0    & 36 & -95.0   & -94.1  & 89   \\
	10 & -20.1   & -20.1 & 0    & 37 & -98.3   & -97.3  & 98   \\
	11 & -22.3   & -22.3 & 99   & 38 & -100.6  & -100.6 & 88   \\
	12 & -25.6   & -25.6 & 96   & 39 & -103.8  & -103.7 & 86   \\
	13 & -27.8   & -27.8 & 21   & 40 & -107.2  & -106.0 & 79   \\
	14 & -31.0   & -31.0 & 95   & 41 & -109.4  & -108.5 & 96   \\
	15 & -33.2   & -33.3 & 78   & 42 & -112.8  & -112.7 & 98   \\
	16 & -36.5   & -36.5 & 5    & 43 & -116.0  & -114.1 & 97   \\
	17 & -38.7   & -38.7 & 79   & 44 & -119.4  & -117.3 & 93   \\
	18 & -42.0   & -42.0 & 77   & 45 & -121.6  & -120.6 & 96   \\
	19 & -45.3   & -45.3 & 87   & 46 & -125.0  & -122.9 & 87   \\
	20 & -47.5   & -47.6 & 76   & 47 & -128.6  & -125.3 & 94   \\
	21 & -50.8   & -50.8 & 89   & 48 & -131.6  & -129.5 & 95   \\
	22 & -53.0   & -53.1 & 99   & 49 & -133.8  & -130.4 & 78   \\
	23 & -56.3   & -56.4 & 74   & 50 & -137.2  & -134.0 & 93   \\
	24 & -59.6   & -59.6 & 90   & 51 & -140.5  & -137.4 & 72   \\
	25 & -61.9   & -61.9 & 91   & 52 & -143.7  & -140.7 & 95   \\
	26 & -65.1   & -65.1 & 22   & 53 & -146.1  & -142.6 & 98   \\
	27 & -68.4   & -67.4 & 97   & 54 & -149.3  & -147.0 & 99   \\
	28 & -70.7   & -70.7 & 53   & 55 & -152.7  & -150.0 & 68   \\
	29 & -74.0   & -74.0 & 77   &    &         &        &      \\

	\hline 
	\end{tabular}
	\caption{Results with unravelling, guided stochastic, and stochastic
	mutation
	operators}
 \end{table}

 \begin{table}[]
	\label {tab:yessa}
	\centering
	\begin{tabular}{cccc|cccc}
	$N$  & Optimal & Found & Gen. & $N$  & Optimal & Found  & Gen. \\
	\hline
	3  & -3.0    & -3     &    0   & 30 & -77.3   & -76.3    &  89  \\
	4  & -5.1    & -5.1   &    0   & 31 & -79.5   & -78.5    &  81  \\
	5  & -7.2    & -7.2   &    23  & 32 & -82.9   & -82.9    &  95  \\
	6  & -9.3    & -9.4   &    90  & 33 & -86.1   & -86.2    &  98  \\
	7  & -12.5   & -12.5  &    90  & 34 & -88.4   & -86.2    &  93  \\
	8  & -14.7   & -14.7  &    93  & 35 & -91.7   & -90.8    &  99  \\
	9  & -16.9   & -16.9  &    2   & 36 & -95.0   & -94.0    &  95  \\
	10 & -20.1   & -20.1  &    1   & 37 & -98.3   & -96.2    &  86  \\
	11 & -22.3   & -22.3  &    53  & 38 & -100.6  & -98.5    &  90  \\
	12 & -25.6   & -25.6  &    80  & 39 & -103.8  & -102.7   &  94  \\
	13 & -27.8   & -27.8  &    24  & 40 & -107.2  & -104.2   &  61  \\
	14 & -31.0   & -31.0  &    56  & 41 & -109.4  & -108.5   &  87  \\
	15 & -33.2   & -33.23 &    73  & 42 & -112.8  & -111.7   &  95  \\
	16 & -36.5   & -36.5  &    15  & 43 & -116.0  & -112.8   &  90  \\
	17 & -38.7   & -38.8  &    72  & 44 & -119.4  & -115.2   &  92  \\
	18 & -42.0   & -42.1  &    93  & 45 & -121.6  & -117.3   &  99  \\
	19 & -45.3   & -45.3  &    94  & 46 & -125.0  & -123.0   &  96  \\
	20 & -47.5   & -47.5  &    86  & 47 & -128.6  & -126.3   &  88  \\
	21 & -50.8   & -50.8  &    22  & 48 & -131.6  & -127.2   &  77  \\
	22 & -53.0   & -53.0  &    61  & 49 & -133.8  & -130.7   &  67  \\
	23 & -56.3   & -56.4  &    92  & 50 & -137.2  & -134.0   &  83  \\
	24 & -59.6   & -58.7  &    76  & 51 & -140.5  & -134.2   &  98  \\
	25 & -61.9   & -61.9  &    70  & 52 & -143.7  & -138.3   &  98  \\
	26 & -65.1   & -65.1  &    3   & 53 & -146.1  & -142.9   &  45  \\
	27 & -68.4   & -67.4  &    79  & 54 & -149.3  & -146.3   &  85  \\
	28 & -70.7   & -70.7  &    67  & 55 & -152.7  & -149.5   &  97  \\
	29 & -74.0   & -73.0  &    90  &    &         &          &       \\

	\hline 
	\end{tabular}
	\caption{Results with unravelling, guided stochastic, and annealed mutation
	operators}
 \end{table}

 \begin{figure}
	\includegraphics[width=\textwidth]{figs/combined1.png}
	\caption{Best Solution Found for N=5 to N=30}
	\label{fig:combined1}
\end{figure}

 \begin{figure}
	\includegraphics[width=\textwidth]{figs/combined2.png}
	\caption{Best Solution Found for N=35 to N=55}
	\label{fig:combined2}
\end{figure}
